# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
options("scipen"=100, "digits" = 2)
```
## What is mean total number of steps taken per day?
```{r}
steps.per.day <- aggregate(FUN = sum, x = activity$steps, by = list(activity$date), na.rm = TRUE)
```

The mean number of steps per day is `r sum(steps.per.day$x, na.rm=TRUE) / 61 `

## What is the average daily activity pattern?
```{r}
plot(steps.per.day, xlab = "Date", ylab = "Steps", main = "Steps Per Day")

hist(steps.per.day$x, xlab = "Steps", main = "Steps Per Day")
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?

```{r}
steps.per.day$day.of.week <- weekdays(as.Date(steps.per.day$Group.1))

weekend <- subset(steps.per.day, day.of.week %in% c("Saturday", "Sunday"))
weekdays <- subset(steps.per.day, ! day.of.week %in%  c("Saturday", "Sunday"))

hist(weekend$x)
hist(weekdays$x)
weekend.mean <- mean(weekend$x)
weekday.mean <- mean(weekdays$x)

```

The mean number of steps during weekdays is `r weekday.mean`, while the mean number of steps on the weekend is `r weekend.mean`.



